A Hexagon Tile Map Algorithm for Displaying Spatial Data

### 2020-10-05 kobakian-cook.Rmd
tassie plots

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, comment = FALSE}
library(knitcitations)
library(kableExtra)
library(RefManageR)
library(sp)
library(sf)
library(sugarbag)
library(absmapsdata)
library(tidyverse)
library(cartogram)
library(ggthemes)
library(cowplot)

knitr::opts_chunk$set(warning = FALSE, message = FALSE)
options("citation_format" = "pandoc")
BibOptions(check.entries = FALSE, style = "markdown")

@@

```{r choro}
states <- absmapsdata::state2011 %>% 
  filter(!state_name_2011 == "Other Territories") %>% 
  rmapshaper::ms_simplify(keep = 0.25, keep_shapes = TRUE)

@@

sa2thyroid_ERP <- sa2thyroid_ERP %>% 
  filter(!is.na(Population)) %>% 
  mutate(SIR = map_chr(p50, aus_colours)) %>% 
  st_as_sf()

aus_ggchoro <- ggplot(sa2thyroid_ERP) + 
  geom_sf(fill = "lightgrey", data = states %>% filter(state_name_2011=="Tasmania")) + 
  geom_sf(aes(fill = SIR)) + 
  scale_fill_identity() +invthm
```

@@

## Make a plot
library(ggplot2)
aus_hexmap <- ggplot() +
  geom_polygon(aes(x = long,  y = lat, group = interaction(state_name_2011, polygon)),
               fill = "darkgrey",  data = fort_state) +
  geom_polygon(aes(x = long,  y = lat,  group = interaction(sa2_name_2011, polygon)),
               fill = "black",  colour = "darkgrey",  size = 0.1, data = fort_aus) +
  geom_polygon(aes(x = long, y = lat, group = hex_id, fill = SIR), data = fort_hex) +
  scale_fill_identity() +
  invthm + coord_equal()
aus_hexmap
```

@@

aus_ggcont <- ggplot(cont) + 
  geom_sf(fill = "darkgrey",  data = states %>% st_transform(., 3112) %>% filter(state_name_2011 == "Tasmania")) +
  geom_sf(fill = "darkgrey",  data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + 
  invthm + guides(fill=FALSE)
aus_ggcont

ggsave(filename = "figs/aus_ggcont.png", plot = aus_ggcont,
       device = "png", bg = "transparent", dpi = 300,  width = 7, height = 6)

@@

```{r}

###############################################################################
# Non - Contiguous Cartograms

ncont <- cartogram_ncont(sa2thyroid_ERP, k = 1.5,
                         weight = "Population") %>% 
                          st_as_sf()

save(ncont, file = "data/ncont.rda")
load("data/ncont.rda")

tas_ggncont <- ggplot(ncont %>% filter(state_name_2011 == "Tasmania")) + 
  geom_sf(fill = "darkgrey",  data = states %>% st_transform(., 3112) %>% filter(state_name_2011 == "Tasmania")) +
  geom_sf(fill = "darkgrey",  data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + 
  invthm + guides(fill=FALSE)
tas_ggncont

@@

aus_ggncont <- ggdraw(xlim = c(0, 1), ylim = c(0, 2)) +
  draw_plot(tas_ggncont, width = 1, height = 2) +
  draw_plot(hobart, x = 0, width = 0.4, height = 0.4)
aus_ggncont

ggsave(filename = "figs/aus_ggncont.png", plot = aus_ggncont,
       device = "png", bg = "transparent", dpi = 300,  width = 7, height = 6)

@@

```{r}

@@

dorl <- sa2thyroid_ERP %>%
  #mutate(pop = (Population/max(Population))*10) %>% # helpful for all of Australia
  cartogram_dorling(., weight = "Population", k = 0.2, m_weight = 1) %>% st_as_sf()
d <- st_bbox(dorl)
aus_ggdorl <- ggplot(dorl) +
  geom_sf(fill = "darkgrey",  data = states %>% st_transform(., 3112) %>% filter(state_name_2011 == "Tasmania")) +
  geom_sf(fill = "darkgrey",  data = sa2thyroid_ERP) + 
  geom_sf(aes(fill = SIR)) + 
  scale_fill_identity()+
  invthm + guides(fill = FALSE)
aus_ggdorl

ggsave(filename = "figs/aus_ggdorl.png", plot = aus_ggdorl,
       device = "png", bg = "transparent", dpi = 300,  width = 7, height = 6)

@@

```{r mapgrid, width = 8, height = 10}
aus_grid <- ggdraw(xlim = c(0, 2), ylim = c(0, 4)) +
  draw_plot(aus_ggchoro, x = 0, y = 2, width = 1, height = 2) +
  draw_plot(aus_ggcont,  x = 1, y = 2, width = 1, height = 2) +
  draw_plot(aus_ggncont, x = 0, y = 0, width = 1, height = 2) +
  draw_plot(aus_ggdorl,  x = 1, y = 0, width = 1, height = 2)

@@

>>>>

```{r defaults, echo=FALSE, message=FALSE, warning=FALSE, comment = FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      comment = FALSE,
                      echo = FALSE,
                      eval = TRUE,
                      out.width = "90%",
                      retina = 3,
                      dev = "png",
                      fig.pos = "center")
```

```{r setup}
library(knitcitations)
library(kableExtra)
library(RefManageR)
options("citation_format" = "pandoc")
BibOptions(check.entries = FALSE, style = "markdown")

library(sp)
library(sf)
library(sugarbag)
# remotes::install_github("wfmackey/absmapsdata")
library(absmapsdata)
library(tidyverse)
library(cartogram)
library(ggthemes)
library(cowplot)

@@

```{r thyroiddata}
states <- absmapsdata::state2011 %>% 
  filter(!state_name_2011 == "Other Territories") %>% 
  rmapshaper::ms_simplify(keep = 0.25, keep_shapes = TRUE)

@@

sa2thyroid_ERP <- sa2thyroid_ERP %>% 
  filter(!is.na(Population)) %>% 
  mutate(SIR = map_chr(p50, aus_colours)) %>% 
  st_as_sf() 
```

```{r choro, eval=FALSE}
aus_ggchoro <- ggplot(sa2thyroid_ERP) + 
  geom_sf(fill = "lightgrey", data = states %>% filter(state_name_2011=="Tasmania")) + 
  geom_sf(aes(fill = SIR)) + 
  scale_fill_identity() +invthm
```

@@

## Make a plot
library(ggplot2)
aus_hexmap <- ggplot() +
  geom_polygon(aes(x = long,  y = lat, group = interaction(state_name_2011, polygon)),
               fill = "darkgrey",  data = fort_state) +
  geom_polygon(aes(x = long,  y = lat,  group = interaction(sa2_name_2011, polygon)),
               fill = "black",  colour = "darkgrey",  size = 0.1, data = fort_aus) +
  geom_polygon(aes(x = long, y = lat, group = hex_id, fill = SIR), data = fort_hex) +
  scale_fill_identity() +
  invthm + coord_equal()
#aus_hexmap
```

@@

aus_ggcont <- ggplot(cont) + 
  geom_sf(fill = "darkgrey",  data = states %>% st_transform(., 3112) %>% filter(state_name_2011 == "Tasmania")) +
  geom_sf(fill = "darkgrey",  data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + 
  invthm + guides(fill=FALSE)
#aus_ggcont

#ggsave(filename = "figs/aus_ggcont.png", plot = aus_ggcont,
#       device = "png", bg = "transparent", dpi = 300,  width = 7, height = 6)

@@

```{r contcart}

###############################################################################
# Non - Contiguous Cartograms
if (!file.exists("data/ncont.rda")) {
ncont <- cartogram_ncont(sa2thyroid_ERP, k = 1.5,
                         weight = "Population") %>% 
                          st_as_sf()

save(ncont, file = "data/ncont.rda")
}

```{r plotcontcart}
load("data/ncont.rda")

tas_ggncont <- ggplot(ncont %>% filter(state_name_2011 == "Tasmania")) + 
  geom_sf(fill = "darkgrey",  data = states %>% st_transform(., 3112) %>% filter(state_name_2011 == "Tasmania")) +
  geom_sf(fill = "darkgrey",  data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + 
  invthm + guides(fill=FALSE)
#tas_ggncont

@@

aus_ggncont <- ggdraw(xlim = c(0, 1), ylim = c(0, 2)) +
  draw_plot(tas_ggncont, width = 1, height = 2) +
  draw_plot(hobart, x = 0, width = 0.4, height = 0.4)
# aus_ggncont

#ggsave(filename = "figs/aus_ggncont.png", plot = aus_ggncont,
#       device = "png", bg = "transparent", dpi = 300,  width = 7, height = 6)

@@

```{r dorl}

@@

dorl <- sa2thyroid_ERP %>%
  #mutate(pop = (Population/max(Population))*10) %>% # helpful for all of Australia
  cartogram_dorling(., weight = "Population", k = 0.2, m_weight = 1) %>% st_as_sf()
d <- st_bbox(dorl)
aus_ggdorl <- ggplot(dorl) +
  geom_sf(fill = "darkgrey",  data = states %>% st_transform(., 3112) %>% filter(state_name_2011 == "Tasmania")) +
  geom_sf(fill = "darkgrey",  data = sa2thyroid_ERP) + 
  geom_sf(aes(fill = SIR)) + 
  scale_fill_identity() +
  invthm + guides(fill = FALSE)
# aus_ggdorl

#ggsave(filename = "figs/aus_ggdorl.png", plot = aus_ggdorl,
#       device = "png", bg = "transparent", dpi = 300,  width = 7, height = 6)

@@

```{r mapgrid, width = 8, height = 10, eval=FALSE}
aus_grid <- ggdraw(xlim = c(0, 2), ylim = c(0, 4)) +
  draw_plot(aus_ggchoro, x = 0, y = 2, width = 1, height = 2) +
  draw_plot(aus_ggcont,  x = 1, y = 2, width = 1, height = 2) +
  draw_plot(aus_ggncont, x = 0, y = 0, width = 1, height = 2) +
  draw_plot(aus_ggdorl,  x = 1, y = 0, width = 1, height = 2)

@@

```{r tas_displays, width = 10, height = 4}
library(gridExtra)
grid.arrange(aus_ggcont, tas_ggncont, aus_ggdorl, ncol=3)
```

### 2020-10-05 kobakian-cook.tex
tassie plots

\begin{Schunk}
\begin{Sinput}
states <- absmapsdata::state2011 %>% 
  filter(!state_name_2011 == "Other Territories") %>% 
  rmapshaper::ms_simplify(keep = 0.25, keep_shapes = TRUE)

sa2 <- absmapsdata::sa22011 %>% 
  filter(state_name_2011 == "Tasmania") %>% 
  filter(!st_is_empty(geometry)) %>% 
  filter(!state_name_2011 == "Other Territories")

# sa2 <- sa2 %>% rmapshaper::ms_simplify(keep = 0.05, keep_shapes = TRUE)

SIR <- read_csv("data/SIR Downloadable Data.csv") %>% 
  filter(SA2_name %in% sa2$sa2_name_2011) %>% 
  dplyr::select(Cancer_name, SA2_name, Sex_name, p50) %>% 
  filter(Cancer_name == "Thyroid", Sex_name == "Females")
ERP <- read_csv("data/ERP.csv") %>%
  filter(REGIONTYPE == "SA2", Time == 2011, Region %in% SIR$SA2_name) %>% 
  dplyr::select(Region, Value)

# Alternative maps
# Join with sa2 sf object
sa2thyroid_ERP <- SIR %>% 
  left_join(sa2, ., by = c("sa2_name_2011" = "SA2_name")) %>%
  left_join(., ERP %>% 
              dplyr::select(Region, 
              Population = Value), by = c("sa2_name_2011"= "Region")) %>% 
  filter(!st_is_empty(geometry))

sa2thyroid_ERP <- sa2thyroid_ERP %>% 
  filter(!is.na(Population)) %>% 
  mutate(SIR = map_chr(p50, aus_colours)) %>% 
  st_as_sf()

aus_ggchoro <- ggplot(sa2thyroid_ERP) + 
  geom_sf(fill = "lightgrey", data = states %>% filter(state_name_2011=="Tasmania")) + 
  geom_sf(aes(fill = SIR)) + 
  scale_fill_identity() +invthm
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
## Create centroids set
centroids <- absmapsdata::sa22011 %>% 
  filter(!st_is_empty(geometry)) %>% 
  filter(state_name_2011 == "Tasmania") %>% 
 # filter(!sa2_name_2011 == "Lord Howe Island") %>% 
  create_centroids(., "sa2_name_2011")

## Create hexagon grid
grid <- create_grid(centroids = centroids,
                    hex_size = 0.2,
                    buffer_dist = 5)

## Allocate polygon centroids to hexagon grid points
hex_allocated <- allocate(
  centroids = centroids,
  hex_grid = grid,
  sf_id = "sa2_name_2011",
  ## same column used in create_centroids
  hex_size = 0.2,
  ## same size used in create_grid
  hex_filter = 10,
  focal_points = capital_cities,
  width = 35,
  verbose = FALSE
)

## Prepare to plot
fort_hex <- fortify_hexagon(data = hex_allocated,
                            sf_id = "sa2_name_2011",
                            hex_size = 0.2) %>% 
            left_join(sa2thyroid_ERP %>% select(sa2_name_2011, SIR, p50))

fort_aus <- sa2thyroid_ERP %>%
 # filter(!sa2_name_2011 == "Lord Howe Island") %>% 
  fortify_sfc()

fort_state <- states %>% 
  filter(state_name_2011 == "Tasmania") %>% 
  fortify_sfc() %>% 
  filter(long < 159)

## Make a plot
library(ggplot2)
aus_hexmap <- ggplot() +
  geom_polygon(aes(x = long,  y = lat, group = interaction(state_name_2011, polygon)),
               fill = "darkgrey",  data = fort_state) +
  geom_polygon(aes(x = long,  y = lat,  group = interaction(sa2_name_2011, polygon)),
               fill = "black",  colour = "darkgrey",  size = 0.1, data = fort_aus) +
  geom_polygon(aes(x = long, y = lat, group = hex_id, fill = SIR), data = fort_hex) +
  scale_fill_identity() +
  invthm + coord_equal()
aus_hexmap
\end{Sinput}

\includegraphics{kobakian-cook_files/figure-latex/fullhexmap-1} \end{Schunk}

\begin{Schunk}
\begin{Sinput}
sa2thyroid_ERP <- st_transform(sa2thyroid_ERP, 3112)
sa2thyroid_ERP <- sa2thyroid_ERP %>% 
  mutate(sva = sqrt(as.numeric(Population)/as.numeric(albers_sqkm)))

###############################################################################
# Contiguous Cartograms

cont <- sa2thyroid_ERP %>% 
  mutate(Population = Population + 1) %>% 
  cartogram_cont(., weight = "Population", itermax = 20) %>%
  st_as_sf()

save(cont, file = "data/cont.rda")
load("data/cont.rda")

aus_ggcont <- ggplot(cont) + 
  geom_sf(fill = "darkgrey",  data = states %>% st_transform(., 3112) %>% filter(state_name_2011 == "Tasmania")) +
  geom_sf(fill = "darkgrey",  data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + 
  invthm + guides(fill=FALSE)
aus_ggcont
\end{Sinput}

\includegraphics{kobakian-cook_files/figure-latex/alternatives-1} \begin{Sinput}
ggsave(filename = "figs/aus_ggcont.png", plot = aus_ggcont,
       device = "png", bg = "transparent", dpi = 300,  width = 7, height = 6)
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
###############################################################################
# Non - Contiguous Cartograms

ncont <- cartogram_ncont(sa2thyroid_ERP, k = 1.5,
                         weight = "Population") %>% 
                          st_as_sf()

save(ncont, file = "data/ncont.rda")
load("data/ncont.rda")

tas_ggncont <- ggplot(ncont %>% filter(state_name_2011 == "Tasmania")) + 
  geom_sf(fill = "darkgrey",  data = states %>% st_transform(., 3112) %>% filter(state_name_2011 == "Tasmania")) +
  geom_sf(fill = "darkgrey",  data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + 
  invthm + guides(fill=FALSE)
tas_ggncont
\end{Sinput}

\includegraphics{kobakian-cook_files/figure-latex/unnamed-chunk-1-1} \begin{Sinput}
# Hobart
hobart <- ggplot(ncont %>% filter(gcc_name_2011 == "Greater Hobart")) + 
  geom_sf(fill = "darkgrey",  data = sa2thyroid_ERP %>% filter(gcc_name_2011 == "Greater Hobart")) +
  geom_sf(aes(fill = SIR, size = 0.001), colour = "black") + 
  scale_fill_identity() + 
  scale_size_identity() +
  invthm + guides(fill=FALSE) + 
  theme(plot.background = element_rect(fill = "transparent", colour = NA)) + 
  theme(panel.background = element_rect(fill = "transparent", colour = "white"))

aus_ggncont <- ggdraw(xlim = c(0, 1), ylim = c(0, 2)) +
  draw_plot(tas_ggncont, width = 1, height = 2) +
  draw_plot(hobart, x = 0, width = 0.4, height = 0.4)
aus_ggncont
\end{Sinput}

\includegraphics{kobakian-cook_files/figure-latex/unnamed-chunk-1-2} \begin{Sinput}
ggsave(filename = "figs/aus_ggncont.png", plot = aus_ggncont,
       device = "png", bg = "transparent", dpi = 300,  width = 7, height = 6)
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
###############################################################################
# Non - Contiguous Dorling Cartograms

dorl <- sa2thyroid_ERP %>%
  #mutate(pop = (Population/max(Population))*10) %>% # helpful for all of Australia
  cartogram_dorling(., weight = "Population", k = 0.2, m_weight = 1) %>% st_as_sf()
d <- st_bbox(dorl)
aus_ggdorl <- ggplot(dorl) +
  geom_sf(fill = "darkgrey",  data = states %>% st_transform(., 3112) %>% filter(state_name_2011 == "Tasmania")) +
  geom_sf(fill = "darkgrey",  data = sa2thyroid_ERP) + 
  geom_sf(aes(fill = SIR)) + 
  scale_fill_identity()+
  invthm + guides(fill = FALSE)
aus_ggdorl
\end{Sinput}

\includegraphics{kobakian-cook_files/figure-latex/unnamed-chunk-2-1} \begin{Sinput}
ggsave(filename = "figs/aus_ggdorl.png", plot = aus_ggdorl,
       device = "png", bg = "transparent", dpi = 300,  width = 7, height = 6)
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
aus_grid <- ggdraw(xlim = c(0, 2), ylim = c(0, 4)) +
  draw_plot(aus_ggchoro, x = 0, y = 2, width = 1, height = 2) +
  draw_plot(aus_ggcont,  x = 1, y = 2, width = 1, height = 2) +
  draw_plot(aus_ggncont, x = 0, y = 0, width = 1, height = 2) +
  draw_plot(aus_ggdorl,  x = 1, y = 0, width = 1, height = 2)

aus_grid
\end{Sinput}

\includegraphics{kobakian-cook_files/figure-latex/mapgrid-1} \begin{Sinput}
ggsave(filename = "figs/aus_grid.png", plot = aus_grid,
  device = "png", bg = "transparent", dpi = 300, width = 8, height = 10)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
aus_hexmap
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
data(capital_cities)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
centroids <- create_centroids(shp_sf = tas_sa2, sf_id = "SA2_NAME16")
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
grid <- create_grid(centroids = centroids, hex_size = 0.2, buffer_dist = 1.2)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
hexmap_allocation <- allocate(
  centroids = centroids %>% select(SA2_NAME16, longitude, latitude),
  sf_id = "SA2_NAME16",
  hex_grid = grid,
  hex_size = 0.2, ## same size used in create_grid
  hex_filter = 10,
  width = 35,
  focal_points = capital_cities,
  verbose = TRUE)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
install.packages("sugarbag")
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
devtools::install_github("srkobakian","sugarbag")
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
library(sugarbag)
\end{Sinput}
\end{Schunk}

>>>>

\begin{Schunk}

\includegraphics[width=0.9\linewidth]{kobakian-cook_files/figure-latex/tas_displays-1} \end{Schunk}

@@

@@

@@

@@

@@

@@

@@

@@
